Lab3--Ping & Kay

Kay is responsible for graph 1 and 2 (i.e., the line and bar chart).
Ping is responsible for graph 3 and 4 (i.e., the heat map and the US map).



```{r}
library(tidyverse)
library(here)
```

```{r}
main_data = read.csv(here("honeyproduction.csv"))
```

First draft for plot_1
```{r}
ggplot(main_data, aes(year, totalprod, group = state))+
  geom_line(aes(color = state))+
  labs (x = "Year",
        y = "Total production of honey",
        title = "The total production of honey of each state from 1998 to 2012")+
  theme_minimal()
 
```

Highlight the west coast states
```{r}
library(gghighlight)
library(colorblindr)
library(colorspace)
```


```{r}
ggplot(main_data, aes(year, totalprod, group = state))+
  geom_line(aes(color = state))+
  gghighlight(state == c("WA", "OR", "CA"))+
  labs (x = "Year",
        y = "Total production of honey",
        title = "The total production of honey of west coast states from 1998 to 2012")+
  theme_minimal()
```

Reproducing three different kinds of color blindness, as well as a desaturated version
```{r}
cvd_grid(ggplot(main_data, aes(year, totalprod, group = state))+
  geom_line(aes(color = state))+
  gghighlight(state == c("WA", "OR", "CA"))+
  labs (x = "Year",
        y = "Total production of honey")+
  theme_minimal())
 
```

Reproduce the plot using a color blind safe palette.
```{r}
ggplot(main_data, aes(year, totalprod, group = state))+
  geom_line(aes(color = state))+
  gghighlight(state == c("WA", "OR", "CA"))+
  scale_color_OkabeIto()+
  labs (x = "Year",
        y = "Total production of honey",
        title = "The total production of honey of west coast states from 1998 to 2012")+
  theme_minimal()
```

Download the region file
```{r}
region <- "https://raw.githubusercontent.com/cphalpert/census-regions/master/us%20census%20bureau%20regions%20and%20divisions.csv"
region_data <- read.csv(region)
```

Join the tables by full_join
```{r}
joined_data <- full_join(main_data, region_data, by = c("state" = "State.Code"))
```

Mutate avg_honey
```{r}
honey_avgprod <-joined_data %>% 
  group_by(state) %>% 
  mutate(avg_honey_bystate = mean(totalprod))
```

Produce a bar plot displaying the average honey for each state collapsed across years.
```{r, fig.width = 10, fig.height= 15}
ggplot(honey_avgprod, aes(fct_reorder(State, avg_honey_bystate), avg_honey_bystate))+
  geom_col()+
  coord_flip()
```

```{r, fig.width = 10, fig.height= 12}
honey_avgprod %>% drop_na(avg_honey_bystate) %>% 
ggplot(aes(fct_reorder(State, avg_honey_bystate), avg_honey_bystate))+
  geom_col(aes(fill = Region))+
  labs(x = "Average amount of honey production",
       y = "State",
       title = "The average honey production of each state from 1998 to 2012")+
  theme_minimal()+
  theme(plot.title = element_text(size = 20))+
  scale_fill_OkabeIto()+
  coord_flip()
```

Create a heatmap displaying the average honey production across years by region

Mutate avgprod_region into the previous dataframe.
```{r}
honey_avgprod_region <-honey_avgprod %>% 
  group_by(year, Region) %>% 
  mutate(avg_honey_byregion = mean(totalprod, na.rm = TRUE))
```

```{r}
ggplot(honey_avgprod_region, aes(year, fct_reorder(Region, avg_honey_byregion, na.rm = TRUE)))+
  geom_tile(aes(fill = avg_honey_byregion), color = "white")+
  scale_fill_viridis_c(option = "magma")+
  labs(x = "Year",
       y = "Region",
       fill = "Average honey \nproduction",
       title = "The average honey production of each region from 1998 to 2012")+
  theme_minimal()+
  theme(plot.title = element_text(size = 15))
```

Another one more plot of interest will be displaying the average honey for each state collapsed across years mapped to the US map.


```{r}
state_data <- map_data("state") %>% 
  rename(State = region)
```

Select only the wanted variables and convert to a lower case to join the data
```{r}
selected_avgprod <- honey_avgprod %>% 
  mutate(State = tolower(State)) %>% 
  select(State, avg_honey_bystate)
```

Delete the duplicate rows in selected_avgprod 
```{r}
unique_selected_avgprod <- unique(selected_avgprod)
```


Join the two dataframes
```{r}
honey_usmap <- left_join(state_data, selected_avgprod)
```

Plot the data to the map.
```{r}
ggplot(honey_usmap) +
  geom_polygon(aes(long, lat, group = group, fill = avg_honey_bystate))+
  coord_fixed()+
  scale_fill_viridis_c()+
   labs(fill = "Average honey production",
       title = "The average honey production of each region from 1998 to 2012")+
   theme(plot.title = element_text(size = 15, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position = "bottom")
```
